Summer 2024 Report

### 2024-11-07 Relations/Report.tex
RP-iffRP

The most interesting work we have carried out in investigating Chapter 1 of TeReSe has resulted from trying to provide constructive proofs for Newman's lemma, the Knaster-Tarski lemma, theorem 1.2.2 and theorem 1.2.3. Before exploring those proofs the following definitions are essential.
\begin{dfn}
  An abstract reduction system (\emph{ARS}) is a structure $\mathcal{A} = (A, \to_\alpha)$ where $A$ is a set and $\to_\alpha$ is a set of binary relations.    
\end{dfn}
A relation $\to$ from the set $\to_\alpha$ is refered to as a reduction (or rewrite) relation. For example, if we specify $\to$ as the relation of being 'less than' ($<$) then $4 \to 2$. \textcolor{red}{Right?}

@@

\begin{dfn}
    A reduction relation is \emph{weakly confluent} (AKA weakly Church-Rosser (WCR)) if all single-step reductions from a given element can always converge via multi-step reductions to some common reduct.
\end{dfn}
\begin{figure}[h]
  \centering
  \begin{tikzpicture}[node distance=1.5cm, auto]
    % Nodes
    \node (top) {a};
    \node (left) [below left of=top] {b};
    \node (right) [below right of=top] {c};
    \node (bottom) [below of=top] {d};

    % Arrows
    \draw[->] (top) -- (left);
    \draw[->] (top) -- (right);
    \draw[->>] (left) -- (bottom);
    \draw[->>] (right) -- (bottom);
\end{tikzpicture}
\caption{An example of a weakly confluent relation.}
\end{figure}

\begin{dfn}
  A reduction relation is \emph{confluent} (AKA Church-Rosser (CR)) if all multi-step redctions from a given element can always converge via multi-step reductions to some common reduct.
\end{dfn}
    \begin{figure}
      \centering
      \begin{tikzpicture}[node distance=1.5cm, auto]
        % Nodes
        \node (top) {a};
        \node (left) [below left of=top] {b};
        \node (right) [below right of=top] {c};
        \node (bottom) [below of=top] {d};

@@

\begin{dfn}
  \emph An element $a \in A$ is a {normal form} if there exists no reduction relation to any other element in $A$.
\end{dfn}

\begin{dfn}
  $a \in A$ is \emph{weakly normalizing} if $a \tto b$ for some normal form $b \in A$. 
  $\to$ is weakly normalizing (WN) if every $a \in A$ is weakly normalizing.  
\end{dfn}
\sacomment{A lot of these are near word for word from TeReSe, is that okay?}

\begin{dfn}
  $a \in A$ is \emph{strongly normalizing} if every reduction sequence starting from $a$ is finite. $\to$ is strongly normalizing (SN) if every 
  $a \in A$ is strongly normalizing. 
\end{dfn}

The above definition of SN is taken from TeReSe. We make use of the alternative definition presented in TeReSe ($\leftarrow$ is accessibly well founded) when working in Agda.

@@

@@

\begin{dfn}
  A \emph{sequence} is a relation from the natural numbers onto a set: $\nat \to A$.
\end{dfn}

\begin{dfn}
  A sequence is \emph{increasing} if every term is greater than (\sacomment{or equal to?}) its previous term.
\end{dfn}

\begin{dfn}
  A sequence is \emph{$\omega$-bounded} if there exists an element $a$ to which all elements of the sequence 
  reduce. 
\end{dfn}

@@

\begin{dfn}
  $a \in A$ has the \emph{recurrence property} if $a \tto b \implies b \tto a$ for all $b \in A$. 
\end{dfn}
\sacomment{I've been using $\tto$ as interchangable with $R *$ (as in the above definition). 
Is that okay or am I going to have to clarify why they are interchangable (are they not interchangable?)?}

\begin{dfn}
  $\to$ has the \emph{recurrence property} (RP) if there is an increasing sequence $f$ in $A$ with an $\omega$-bound and 
  there is an element of the sequence that has the recurrence property. 
\end{dfn}

@@

        \end{figure}
\begin{dfn}
  $\to$ has the \emph{recurrence property -} (RP-)  if there is an increasing sequence $f$ in $A$ with an $\omega$-bound $a$ and 
  there is an element $i$ of the sequence such that $a \tto f (i)$.
\end{dfn}

>>>>

The most interesting work we have carried out in investigating Chapter 1 of TeReSe has resulted from trying to provide constructive proofs for Newman's lemma, the Knaster-Tarski lemma, theorem 1.2.2 and theorem 1.2.3. Before exploring those proofs the following definitions are essential.

The following definitions are taken from [TeReSe], Chapter 1,
with minimal adaptations for consistency.

\begin{dfn}
  An abstract reduction system (\emph{ARS}) is a structure $\mathcal{A} = (A, \to_\alpha)$ where $A$ is a set and $\to_\alpha$ is a set of binary relations.
\end{dfn}
A relation $\to$ from the set $\to_\alpha$ is refered to as a reduction (or rewrite) relation. For example, if we specify $\to$ as the relation of being 'less than' ($<$) then $4 \to 2$. \textcolor{red}{Right?}
\apcomment{Yes, but move this to a(n eventual) list of examples.}

@@

\apcomment{Thus, when $(a,b) \in \to_{(\alpha)}$, we say that there is a
\emph{$(\alpha-)$reduction} from $a$ to $b$.
When $(a,b) \in \tto_\alpha$, we say that there is a multi-step $\alpha$-reduction
from $a$ to $b$.}

\begin{dfn}
    A reduction relation is \emph{weakly confluent} (AKA weakly Church-Rosser (WCR)) if all single-step reductions from a given element can always converge via multi-step reductions to some common reduct.
\end{dfn}
\begin{figure}[h]
  \centering
\[  \begin{tikzpicture}[node distance=1.5cm, auto]
    % Nodes
    \node (top) {a};
    \node (left) [below left of=top] {b};
    \node (right) [below right of=top] {c};
    \node (bottom) [below right of=left] {d};

    % Arrows
    \draw[->] (top) -- (left);
    \draw[->] (top) -- (right);
    \draw[->>] (left) -- (bottom);
    \draw[->>] (right) -- (bottom);
\end{tikzpicture}
\hspace{3cm}
\begin{tikzpicture}[node distance=1.5cm, auto]
  % Nodes
  \node (top) {a};
  \node (left) [below left of=top] {b};
  \node (right) [below right of=top] {c};
  \node (bottom) [below right of=left] {d};

  % Arrows
  \draw[->>] (top) -- (left);
  \draw[->>] (top) -- (right);
  \draw[->>] (left) -- (bottom);
  \draw[->>] (right) -- (bottom);
\end{tikzpicture}
\]
\caption{An example of a weakly confluent relation.}
\end{figure}

\begin{dfn}
  A reduction relation is \emph{confluent} (AKA Church-Rosser (CR)) if all multi-step reductions from a given element can always converge via multi-step reductions to some common reduct.
\end{dfn}
    \begin{figure}
      \centering
      \begin{tikzpicture}[node distance=1.5cm, auto]
        % Nodes
        \node (top) {a};
        \node (left) [below left of=top] {b};
        \node (right) [below right of=top] {c};
        \node (bottom) [below right of=left] {d};

@@

\begin{dfn}
  \emph An element $a \in A$ is a {normal form} if there exists no reduction to any other element in $A$.
\end{dfn}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-weakly normalizing} ($WN_{R}$)if $a \tto b$ for some normal form $b \in A$.

  A relation $\to$ is weakly normalizing (WN) if every $a \in A$ is $R$- weakly normalizing.
\end{dfn}
\sacomment{A lot of these are near word for word from TeReSe, is that okay?}

\begin{dfn}
  An element $a \in A$ is \emph{$R$-strongly normalizing} ($SN_R$) if every reduction sequence starting from $a$ is finite.

  A relation $\to$ is strongly normalizing (SN) if every
  $a \in A$ is $\to$-strongly normalizing.
\end{dfn}

The above definition of SN is taken from TeReSe. We make use of the alternative definition presented in TeReSe ($\leftarrow$ is accessibly well-founded) when working in Agda.

@@

\apcomment{This is not quite right; rather, $\to^=$ is the reflexive-symmetric-transitive closure of $\to$, i.e., it is the \emph{equivalence relation generated by $\to$}, the NFP refers to this latter relation.}

\apcomment{Add a subsection/definition listing the closure operations on relations.}

@@

Let $A$ be a set.
\begin{dfn}
  A \emph{sequence (in $A$)} is a function from the natural numbers to $A$:
  \begin{align*}
    &s : \nat \to A \\
    &s = (s_0,s_1,s_2,\dots)
  \end{align*}
\end{dfn}

\begin{dfn}
  A sequence is \emph{$R$-increasing} if every term is $R$-related to its preceding term.
\end{dfn}

\begin{dfn}
  A sequence is \emph{bounded} if there exists an element $a$ to which all elements of the sequence reduce.
\end{dfn}

@@

\begin{dfn}
  $a \in A$ has the \emph{recurrence property} if $a \tto b \implies b \tto a$ for all $b \in A$.
\end{dfn}
\sacomment{I've been using $\tto$ as interchangable with $R *$ (as in the above definition).
Is that okay or am I going to have to clarify why they are interchangable (are they not interchangable?)?}
\apcomment{It's okay, but later we will need to make sure we are using them consistently.  Say that in Agda code, $\to$ is denoted by $R$.}

\begin{dfn}
  $\to$ has the \emph{recurrence property} (RP) if every
  bounded, increasing sequence has a recurrent element.
\end{dfn}

@@

        \end{figure}
\begin{dfn}
  $\to$ has the \emph{weak recurrence property} (RP-) if
  for every increasing sequence with bound $a$,
  there is an element $i$ of the sequence such that $a \tto f (i)$.
\end{dfn}

### 2024-11-07 Relations/notes.tex
RP-iffRP

>>>>

\section{Todo}
\begin{enumerate}
  \item In the report, give a classical proof that the $Inc$ Property
  implies that no $R$-increasing sequence is bounded.

  Conclude that every relation satisfying $Inc$ also satisfies $RP$.

  Therefore, the proof of Theorem 1.2.3(iii) using $RP$ instead of $Inc$
  is generalizing the claim of the theorem (as stated in [TeReSe]).
\end{enumerate}
